\ProjectName{}: Data Imputation as a Query Optimization

### 2017-02-22 paper/algorithm.tex
paper: cleaned up complexity section, removed todos

\subsection{Complexity}
\todo{this section needs to be cleaned up/clarified}
Our optimization algorithm builds off the approach taken by System R \cite{blasgen1981system}, therefore our algorithm still operates in exponential time. Indeed, 
note that if we remove our restriction on imputation operators (i.e. minimal vs maximal), and allow any arbitrary subset of attributes to be imputed,
then every single operator in the query plan has a number of imputations exponential in the number of dirty columns. Our restriction, instead, increases the number
of plans (in the worst case) at each operator by a factor of 3. Of course, this implies that in the worst case (where all dirty sets tracked are distinct throughout the query plan),
we explore a number of plans that further scales the number of plans in the original algorithm by an exponential factor.

However, we note that in all cases we have considered, this exponential blowup does not affect the practical performance of our optimizer.

>>>>

\subsection{Complexity}
Our optimization algorithm builds off the approach taken by System R \cite{blasgen1981system}, therefore our planning algorithm still operates in exponential time. 
If we removed our restriction on imputation operators to consider not just minimal versus maximal imputations, and instead allowed any arbitrary subset
of attributes to be imputed at a plan node (including those not operated on by that node), then every single operator in the query plan would have a
number of plans exponential in the number of dirty columns.

By restricting imputations to two classes: those that impute only attributes used in the operator and those that impute all attributes needed downstream in the query,
we limit the factor of plans at each operator to three: \textit{Drop}, \textit{Impute} over attributes used in operator, and \textit{Impute} over all attributes needed.
Recall that the planner maintains different plans for different dirty sets, keeping only those that are not dominated by other plans with the same dirty set. So in many
cases we can drop some of these intermediate plans generated at each operator. However, if the dirty sets tracked are distinct through the entire planning phase,
then we could explore $\mathcal{O}(3^{\# of operators}) * plans$.

However, we note that in all cases we have considered, this exponential blowup does not affect the practical performance of our optimizer and results in relatively short
planning times.

### 2017-02-22 paper/experiments.tex
paper: cleaned up complexity section, removed todos

\subsection{Queries}
\todo{Add more queries: we want ~10 (balanced between the two datasets), explain them below}
We collected a set of queries (Table~\ref{fig:queries}) that we think are both representative, in that they could reasonably be written by a user in the course of data analysis, and interesting to plan.

>>>>

\subsection{Queries}
We collected a set of queries (Table~\ref{fig:queries}) that we think are both representative, in that they could reasonably be written by a user in the course of data analysis, and interesting to plan.
